cardinalR: Generating interesting high-dimensional data structures

### 2025-04-20 paper-cardinalR.Rmd
Updated first intro paragraph

The generation of synthetic datasets with well-defined structural properties is a cornerstone of evaluating and benchmarking algorithms across diverse fields, including machine learning, data mining, and computational biology. Researchers frequently require the ability to create data with specific dimensionalities, noise characteristics, and intricate underlying structures to rigorously test the performance and robustness of their methods. The R ecosystem offers a variety of packages for generating synthetic data, each with its strengths and focus. For instance, **`geozoo` (Schloerke, [Year of last commit on GitHub])** provides a rich collection of geometric objects, facilitating the creation and analysis of specific shapes, primarily in lower-dimensional spaces. **`snedata` (Melville, [Year of last commit on GitHub])** offers tools for generating simplified datasets tailored for evaluating dimensionality reduction techniques like t-SNE, often focusing on revealing low-dimensional embeddings of complex structures. Packages like **`mlbench` (Leisch & Dimitriadou, 2010)** provide a collection of well-known benchmark datasets, often with established classification or regression challenges. In the realm of single-cell omics, **`splatter` (Zappia et al., 2017)** excels at simulating complex biological data, capturing nuances like batch effects and differential expression.

>>>>

Generating synthetic datasets with clearly defined geometric properties is essential for evaluating and benchmarking algorithms in various fields, such as machine learning, data mining, and computational biology. Researchers often need to generate data with specific dimensions, noise characteristics, and complex underlying structures to test the performance and robustness of their methods.

The R ecosystem offers a variety of packages for generating synthetic data, each with its strengths and focus. For instance, **`geozoo` (Schloerke, [Year of last commit on GitHub])** provides a rich collection of geometric objects, facilitating the creation and analysis of specific shapes, primarily in lower-dimensional spaces. **`snedata` (Melville, [Year of last commit on GitHub])** offers tools for generating simplified datasets tailored for evaluating dimensionality reduction techniques like t-SNE, often focusing on revealing low-dimensional embeddings of complex structures. Packages like **`mlbench` (Leisch & Dimitriadou, 2010)** provide a collection of well-known benchmark datasets, often with established classification or regression challenges. In the realm of single-cell omics, **`splatter` (Zappia et al., 2017)** excels at simulating complex biological data, capturing nuances like batch effects and differential expression.
